---
title: "Homework 1"
author: "Nathan Yang"
date: "8-30-2024"
output:
  pdf_document: default
---

### Exercise 1

a.  Load the Data using the read.table() function

    ```{r part-a}
    rain_df <- read.table(file = "../homeworks/homework-1/data/rnf6080.dat")
    ```

b.  I can use `dim(rain_df)` to get the number of rows and columns

    ```{r part-b}
    dim(rain_df)
    ```

    From this, I know that `rain_df` has 5070 rows and 27 columns

c.  I can use `colnames(rain_df)` to get all the column names

    ```{r part-c}
    colnames(rain_df)
    ```

d.  I can index the dataframe to get the value in the 2nd row 4th column

    ```{r part-d}
    rain_df[2, 4]
    ```

e.  I can display the whole second row by indexing it with the column value removed

    ```{r part-e}
    rain_df[2,]
    ```

f.  This command renames the column names of rain_df to be "year", "month", "day", and the numbers 0 through 23

    ```{r}
    names(rain_df) <- c("year","month","day",seq(0,23))
    ```

g.  Adding a column that aggregates daily rainfall

    ```{r message=TRUE, warning=TRUE}
    library(dplyr)
    rain_df <- rain_df |>
      mutate(
        daily_rain_fall = rowSums(select(rain_df, one_of(as.character(0:23))))
      )
    ```

h.  Creating a histogram

    ```{r}
    library(ggplot2)
    ggplot(rain_df) +
      geom_col(
        aes(
          x = as.Date(paste(year, month, day, sep = "-")), 
          y = daily_rain_fall
          )
        ) +
      labs(x = "Date", y = "Rainfall", title = "Daily Plot of Rainfall")
    ```

i.  This histogram can't be right because there can't exist days with negative rainfall

j.  Here is how I would "fix" the dataframe.\
    First, I needed to look at the rows that had negative values for `daily_rain_fall`

    ```{r}
    weird_rows <- rain_df |> 
      filter(daily_rain_fall < 0)
    tibble(weird_rows)
    ```

    Next, I need to replace all the all the "-999" values since they are obviously incorrect. Here I chose to replace them all with 0s

    ```{r}
    imputed_df <- rain_df |>
      mutate_all(~ replace(., . == -999, 0))

    ```

    Next I would need to recalculate the daily rainfall sums

    ```{r}
    corrected_df <- imputed_df |>
      mutate(
        daily_rain_fall = rowSums(select(imputed_df, one_of(as.character(0:23))))
      )
    ```

k.  Here is the regenerated histogram

    ```{r}
    ggplot(corrected_df) +
      geom_col(
        aes(
          x = as.Date(paste(year, month, day, sep = "-")), 
          y = daily_rain_fall
          )
        ) +
      labs(x = "Date", y = "Rainfall", title = "Daily Plot of Rainfall")
    ```

    This is more reasonable than the previous histogram because there are no more negative values and the range of values is much more in line with expected rainfall.

### Exercise 2

a.  The result looks strange because the numbers are treated as strings here.\
    `max(x)` returns "7" due to it being the latest alphanumerically\
    `sort(x)` returns `["12", "5", "7"]` as this is the alphanumeric order\
    `sum(x)` is an error since strings cannot be added together by `sum()`

    ```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
    x <- c("5", "12", "7")
    max(x)
    sort(x)
    sum(x)
    ```

b.  The `+` operation produces an error because one variable being a string determined `y` to be a vector of strings

    ```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
    y <- c("5", 7, 12)
    y[2] + y[3]
    ```

c.  The `+` operation works correctly because the 2 referenced values were numbers. Only the first value was determined to be a string while the second and third values were added.

    ```{r}
    z <- data.frame(z1 = "5", z2 = 7, z3 = 12)
    z[1,2] + z[1,3]
    ```

### Exercise 3

a.  The point of reproducible code is to ensure other people are able to confirm my work/findings as well as demonstrate a complete workflow.

b.  An example of why reproducible code is so important is to verify my methods were correct, consistent, and statistically acceptable.

c.  I rate this assignment in terms of difficulty a 6/10 since I struggled with correcting and aggregating the daily rainfall
